---
title: "hw02-Tyler-Larsen"
output: github_document
---

```{r}
library(readr)
library(ggplot2)
library(dplyr)
```
##1. CREATE DATA DICTIONARY FILE
* Refer to Github

##2. IMPORT THE DATA IN R
```{r}
dat <- read.csv('nba2017-player-statistics.csv', stringsAsFactors = FALSE)
str(dat)

dat$Player <- as.character(dat$Player)
dat$Team <- as.character(dat$Team)
dat$Experience <- as.character(dat$Experience)
dat$Position <- as.factor(dat$Position)
dat$Salary <- as.double(dat$Salary)

dat$Rank <- as.integer(dat$Rank)
dat$Age <- as.integer(dat$Age)
dat$GP <- as.integer(dat$GP)
dat$GS <- as.integer(dat$GS)
dat$MIN <- as.integer(dat$MIN)
dat$FGM <- as.integer(dat$FGM)
dat$Points3 <- as.integer(dat$Points3)
dat$Points3_atts <- as.integer(dat$Points3_atts)
dat$Points2 <- as.integer(dat$Points2)
dat$Points2_atts <- as.integer(dat$Points2_atts)
dat$PointsFTM <- as.integer(dat$FTM)
dat$FTA <- as.integer(dat$FTA)
dat$OREB <- as.integer(dat$OREB)
dat$DREB <- as.integer(dat$DREB)
dat$AST <- as.integer(dat$AST)
dat$STL <- as.integer(dat$STL)
dat$BLK <- as.integer(dat$BLK)
dat$TO <- as.integer(dat$TO)

```

```{r}
dat2 <- read_csv('nba2017-player-statistics.csv')
str(dat2)

dat2$Player <- as.character(dat2$Player)
dat2$Team <- as.character(dat2$Team)
dat2$Experience <- as.character(dat2$Experience)
dat2$Position <- as.factor(dat2$Position)
dat2$Salary <- as.double(dat2$Salary)

dat2$Rank <- as.integer(dat2$Rank)
dat2$Age <- as.integer(dat2$Age)
dat2$GP <- as.integer(dat2$GP)
dat2$GS <- as.integer(dat2$GS)
dat2$MIN <- as.integer(dat2$MIN)
dat2$FGM <- as.integer(dat2$FGM)
dat2$Points3 <- as.integer(dat2$Points3)
dat2$Points3_atts <- as.integer(dat2$Points3_atts)
dat2$Points2 <- as.integer(dat2$Points2)
dat2$Points2_atts <- as.integer(dat2$Points2_atts)
dat2$PointsFTM <- as.integer(dat2$FTM)
dat2$FTA <- as.integer(dat2$FTA)
dat2$OREB <- as.integer(dat2$OREB)
dat2$DREB <- as.integer(dat2$DREB)
dat2$AST <- as.integer(dat2$AST)
dat2$STL <- as.integer(dat2$STL)
dat2$BLK <- as.integer(dat2$BLK)
dat2$TO <- as.integer(dat2$TO)
```

##3. RIGHT AFTER IMPORTING THE DATA
```{r}
#subset(dat, Experience == 'R') <- 
```

##4. PERFORMANCE OF PLAYERS
```{r}
#additional variables needed to compute EFF
dat$missed_FG <- dat$FGA - dat$FGM
dat$missed_FT <- dat$FTA - dat$FTM
dat$PTS <- 2*dat$Points2 + 3*dat$Points3 + dat$FTM
dat$REB <- dat$DREB + dat$OREB
dat$MPG <- dat$MIN / dat$GP

#EFF
dat$EFF <- (dat$PTS + dat$REB + dat$AST + dat$STL + dat$BLK - dat$missed_FG - dat$missed_FT - dat$TO) / dat$GP
summary(dat$EFF)

top_EEF <- data.frame("Player" = dat$Player, "Team" = dat$Team, "Salary" = dat$Salary, "EFF" = dat$EFF)
top_EEF <- top_EEF[order(top_EEF$EFF, decreasing = TRUE), ]
head(top_EEF, 10)

subset(dat, dat$EFF < 0)

pts_cor <- cor(dat$EFF, dat$PTS)
reb_cor <- cor(dat$EFF, dat$REB)
ast_cor <- cor(dat$EFF, dat$AST)
stl_cor <- cor(dat$EFF, dat$STL)
blk_cor <- cor(dat$EFF, dat$BLK)
missed_fg_cor <- -1 * cor(dat$EFF, dat$missed_FG)
missed_ft_cor <- -1 * cor(dat$EFF, dat$missed_FT)
to_cor <- -1 * cor(dat$EFF, dat$TO)

EFF_cor <- c(pts_cor, reb_cor, ast_cor, stl_cor, blk_cor, missed_fg_cor, missed_ft_cor, to_cor)
EFF_cor <- sort(EFF_cor, decreasing = TRUE)
EFF_cor
barplot(EFF_cor, col = "red")

```

##5. EFFICIENCY AND SALARY
```{r}
ggplot(dat, aes(x = dat$EFF, y = dat$Salary))+
  geom_point(aes(color = dat$Team))+
  geom_smooth(method = loess)

players2 <- subset(dat, dat$Experience > 19)
head(players2)

ggplot(players2, aes(x = players2$EFF, y = players2$Salary))+
  geom_point(aes(color = players2$Team))+
  geom_smooth(method = loess)

EFF_sal_cor <- cor(players2$EFF,players2$Salary)
```
What can you say about the relationship between these two variables for the set of
“more established players”?
* 

##6.COMMENT AND REFLECTIONS  
* What things were hard, even though you saw them in class/lab?
* What was easy(-ish) even though we haven’t done it in class/lab?
* Did you need help to complete the assignment? If so, what kind of help?
* How much time did it take to complete this HW?
* What was the most time consuming part?
* Was there anything that you did not understand? or fully grasped?
* Was there anything frustrating in particular?
* Was there anything exciting? Something that you feel proud of? (Don’t be shy, we
won’t tell anyone).

















