---
title: "HW 1: More Vectors"
subtitle: "Stat 133, Fall 2017"
author: "Gaston Sanchez"
output: github_document
urlcolor: blue
---
  
```{r}
#laod data file
load("data/nba2017-salary-points.Rdata")

#view data file contents
ls()
```

# Research Question:
"The more points a player scores, the higher the salary?"


## 1. A Bit of Data Processing  

```{r}
#mak salary in terms of millions
salary_mil <- salary/1000000
round(salary_mil, 2)

#convert experience into integers
experience_int <- experience
experience_int[experience_int=="R"] <- 0
experience_int <- strtoi(experience_int, 10)
is.integer(experience_int)
experience_int

#convert position to factor, more informative data labels
fposition <- position
fposition[fposition=="C"] <- "center"
fposition[fposition=="SF"] <- "small_fwd"
fposition[fposition=="PF"] <- "power_fwd"
fposition[fposition=="SG"] <- "shooting_guard"
fposition[fposition=="PG"] <- "point_guard"
fposition <- factor(fposition)
table(fposition)
```

## 2. Scatterplot of Points and Salary

```{r}
plot(points, salary_mil, xlab = "Points", ylab = "Salary(millions)", col = "red", cex = .5, main = "Scatterplot of Points and Salary")
```
## 3. Correlation Between Points and Salary

```{r}
#creating the necesary variables for our calculations
n <- length(player)

#points variables
points_mean <- 1/n * sum(points)
round(points_mean, 2)
points_var <- sum((points - points_mean)^2)/(n-1)
points_sd <- sqrt(points_var)

#salary variables
salary_mean <- 1/n * sum(salary_mil)
round(salary_mean, 2)
salary_var <- sum((salary_mil - salary_mean)^2)/(n-1)
salary_sd <- sqrt(salary_var)

sp_cov <- (sum((salary_mil - salary_mean)*(points - points_mean)))/(n-1)
sp_cov
sp_cor <- sp_cov / (salary_sd * points_sd)
sp_cor
```

## 4. Simple Linear Regression

```{r}
slope <- sp_cor * salary_sd/points_sd
slope
intercept<- salary_mean - (points_mean * slope)
intercept
y <- intercept + (slope * points)

summary(y)
```

* Regression Equation: y = `r slope`x + `r intercept`
* The slope coefficent is an indicator of how much a player's salary increases as he scores more points
* The intercept is an indicator of what a player's expected salary is if he scores 0 points
* Predicted Salaries:
    + 0 points:  = $`r round(slope*0 + intercept, 2)` million
    + 100 points:  = $`r round(slope*100 + intercept, 2)` million
    + 500 points:  = $`r round(slope*500 + intercept, 2)` million
    + 1000 points:  = $`r round(slope*1000 + intercept, 2)` million
    + 2000 points:  = $`r round(slope*2000 + intercept, 2)` million

## 5. Plotting the Regression Line

```{r}
plot (points, salary_mil, abline(intercept, slope))
lines(lowess(points,salary_mil), col="blue")

```

##7.Exploring Position and Experience

```{r}
plot(experience_int, salary_mil)
lines(lowess(experience_int,salary_mil), col="blue", text = "Lowess")
# scatterplot3d(points, experience_int, salary_mil)
#boxplot(salary_mil, position)
```





## 8. Comments and Reflections

















